---
output:
  html_document: default
  pdf_document: default
---

```{r message=FALSE, warning=FALSE, include=FALSE}

library(dplyr)
library(tidyr)
library(stringr)
library(synapser)
library(data.table)
library(ggplot2)
library(plotly)

synLogin()

mut<-read.table(synGet("syn5839666")$path, sep = "\t", header = TRUE, colClasses = "character", stringsAsFactors = TRUE)

genelist.1027 <- read.csv(synGet("syn12179299")$path, header = F)

mut$chr <- str_extract(mut$Chromosome, "\\d")
names(mut)[names(mut)=="Mutation_Type"] <- "Variant_Classification"
names(mut)[names(mut)=="Sample_ID"] <- "Tumor_Sample_Barcode"

mut$cosmic[mut$Hugo_Symbol %in% cosmic$Gene.Symbol] <- "T"
mut$cosmic[!mut$Hugo_Symbol %in% cosmic$Gene.Symbol] <- "F"

soms <- filter(mut, Mutation_Status == "Somatic") %>% filter(Hugo_Symbol %in% genelist.1027$V1)

germs <- filter(mut, Mutation_Status == "Germline") %>% filter(Hugo_Symbol %in% genelist.1027$V1)

```

# Mutation Summaries
## Somatic Mutations

####Plot indicates summary count of germline cNF patient mutations detected in 1027 gene list, ignoring "silent", "intron" and "in frame" mutations.

```{r echo=FALSE, fig.width=8, message=FALSE, warning=FALSE}
library(DT)

germs.summary <- germs %>%
  filter(Variant_Classification != "Silent" & Variant_Classification != "Intron" & Variant_Classification != "In_Frame_Ins" &
           Variant_Classification != "In_Frame_Del") %>% 
  select(Tumor_Sample_Barcode, Hugo_Symbol) %>% 
  distinct() %>% 
  group_by(Hugo_Symbol) %>% 
  add_count() %>% 
  select(Hugo_Symbol, n) %>% 
  distinct()
  
p<-ggplot(germs.summary) +
  geom_bar(aes(x=reorder(Hugo_Symbol, -n), y = n, fill = n), stat = "identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Gene", y = "Count")

ggplotly(p)
```

#### Sortable table indicating all germline mutations observed in patient samples, filtered to 1027 gene list, but including "silent", "intron" and "in frame" mutations. 
```{r echo=FALSE, message=FALSE, warning=FALSE}


datatable(germs)
```


## Somatic Mutations

####Plot indicates summary count of somatic cNF mutations detected in 1027 gene list, ignoring "silent", "intron" and "in frame" mutations.

```{r echo=FALSE, fig.width=8, message=FALSE, warning=FALSE}

soms.summary <- soms %>%
  filter(Variant_Classification != "Silent" & Variant_Classification != "Intron" & Variant_Classification != "In_Frame_Ins" &
           Variant_Classification != "In_Frame_Del") %>% 
  select(Tumor_Sample_Barcode, Hugo_Symbol) %>% 
  distinct() %>% 
  group_by(Hugo_Symbol) %>% 
  add_count() %>% 
  select(Hugo_Symbol, n) %>% 
  distinct()
  
p<-ggplot(soms.summary) +
  geom_bar(aes(x=reorder(Hugo_Symbol, -n), y = n, fill = n), stat = "identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Gene", y = "Count")

ggplotly(p)
```

#### Sortable table indicating all somatic mutations observed in cNF samples, filtered to 1027 gene list but including "silent", "intron" and "in frame" mutations.  
```{r echo=FALSE, message=FALSE, warning=FALSE}
datatable(soms)
```

```{r synapse message=FALSE, warning=FALSE, include=FALSE}
this.file <- ""
  
write.csv(soms,"somatic_mutations_1027.txt")
write.csv(germs,"germline_mutations_1027.txt")

synStore(File("somatic_mutations_1027.txt",parentId="syn12179946"), used = c("syn5839666", "syn12179299"), executed = this.file)
synStore(File("germline_mutations_1027.txt",parentId="syn12179946"), used = c("syn5839666", "syn12179299"), executed = this.file)

germ.annot <- list("resourceType" = "analysis", 
              "analysisType" = "variant calling",
              "tumorType" = "Neurofibroma Cutaneous",
              "assay" = "wholeGenomeSeq",
              "diagnosis" = "Neurofibromatosis 1")

som.annot <- list("resourceType" = "analysis", 
              "analysisType" = "somatic mutation calling",
              "tumorType" = "Neurofibroma Cutaneous",
              "assay" = "wholeGenomeSeq",
              "diagnosis" = "Neurofibromatosis 1")

synSetAnnotations("syn12179947", som.annot)
synSetAnnotations("syn12179949", germ.annot)

```
